\section{Gini-Frontier Certification Framework}
Suppose a verifier-gated incumbent has objective $\cutoff$.  Any improving
solution satisfies $G<\cutoff$ because $P\geq0$.  The algorithm partitions the
relevant improving Gini domain into closed intervals
$\Gamma_q=[\gamma_q^L,\gamma_q^U]$ whose interiors are disjoint and whose
union covers that domain.

For each interval, a valid non-enumerative relaxation first supplies a lower
bound.  An interval not fathomed by that bound is passed to the original
fixed-interval compact model.  Its no-improver form includes
\begin{align}
 \gamma_q^L &\leq G\leq\gamma_q^U,\label{eq:ginterval}\\
 G+\lambda P &\leq \cutoff-\eps.\label{eq:cutoff}
\end{align}
CPLEX then proves infeasibility, proves a valid best bound at least
$\cutoff-\eps$, or returns an unresolved leaf with its valid bound.  A timeout
or wrapper checkpoint never becomes an exact closure by itself.

\begin{proposition}[Fixed-interval no-improver validity]
If the model consisting of the original compact formulation,
\eqref{eq:ginterval}, \eqref{eq:cutoff}, and globally valid strengthening
rows is infeasible, then no original feasible solution in $\Gamma_q$ improves
the incumbent.
\end{proposition}
\begin{proof}
Every original improving solution in $\Gamma_q$ satisfies the original model,
the interval rows, and \eqref{eq:cutoff}.  Every strengthening row is a
necessary condition for that same solution.  Such a solution would therefore
be feasible in the fixed-interval model, contradicting infeasibility.
\end{proof}

\begin{theorem}[Full-frontier global exactness]
Let $\widehat x\in\mathcal X$ be independently verified with objective
$\cutoff$.  Assume (i) the final Gini leaves exactly cover every possible
improving Gini value; (ii) each final leaf is closed by a valid relaxation
bound, exact fixed-interval infeasibility, a valid fixed-interval best bound,
or an empty-domain proof; and (iii) no invalid, unresolved, or open leaf
remains.  Then $\widehat x$ is globally optimal.
\end{theorem}
\begin{proof}
If a better solution existed, its Gini value would belong to one final leaf.
The corresponding leaf closure proves that no improving original solution
exists in that leaf, a contradiction.  Hence no feasible solution has
objective below $\cutoff$.
\end{proof}

An optional paper-safe $S$ refinement may partition a fixed Gini leaf's valid
domain $[S^L,S^U]$ into children.  This option was disabled in the tested
route profile.
\begin{proposition}[Optional $S$-ledger disjunction]
If $[S^L,S^U]=\bigcup_b[S_b^L,S_b^U]$ with disjoint interiors and every child
is exactly closed using the enforced child bounds, then the parent Gini leaf
is closed.
\end{proposition}
\begin{proof}
Every parent-feasible point has one denominator value $S$ and therefore lies
in at least one child.  Closure of every child excludes every parent-feasible
improver.  Exact coverage prevents an omitted denominator region.
\end{proof}

\begin{center}
\begin{minipage}{0.9\linewidth}
\textbf{Stable algorithm.} Generate and verify a same-run heuristic incumbent;
construct the full Gini ledger; screen leaves with valid relaxation bounds;
solve unresolved leaves with the tailored compact model and the selected
route-cutset callback; merge only exact leaf evidence; certify only after the
coverage and source audits pass.
\end{minipage}
\end{center}

Operationally, the stable implementation executes the following generic
sequence for every instance:
\begin{enumerate}
\item construct and independently verify a same-run HGA--TGBC incumbent;
\item partition the complete improving Gini domain and initialize its ledger;
\item apply valid non-enumerative bounds and domain propagation to each leaf;
\item solve every unresolved leaf with the original fixed-interval compact
model, paper-safe static rows, and the bounded route-cutset separator;
\item merge only scope-valid leaf bounds or exact closures; and
\item issue a global certificate only after coverage, verifier, source, and
finalization audits all pass.
\end{enumerate}
