\section{CPLEX-Managed Tailored Branch-and-Cut}
The fixed-interval compact model is solved through the CPLEX C API
\cite{cplex2211}.  CPLEX
manages presolve, LP relaxations, branching, node selection, incumbents, best
bounds, and MIP termination.  The tailored layer constructs the original
fixed-interval model, applies the static valid rows of the previous section,
and registers a bounded user-cut separator.

\subsection{Selected directed route-cutset separator}
For vehicle $k$, a station subset $A\subseteq\I$ not containing the depot, and
a representative $\ell\in A$, define the directed boundary
\begin{equation}
 \delta_k(A)=
 \sum_{i\notin A,j\in A}x_{kij}
 +\sum_{i\in A,j\notin A}x_{kij}.
\end{equation}
The callback separates
\begin{equation}
 \delta_k(A)\geq2z_{k\ell}. \label{eq:route-cutset}
\end{equation}

\begin{proposition}[Validity of the callback route cut]
Inequality \eqref{eq:route-cutset} is valid for every original integer route,
regardless of how $A$ was selected.
\end{proposition}
\begin{proof}
If $z_{k\ell}=0$, the right-hand side is zero and the inequality follows from
nonnegativity.  If $z_{k\ell}=1$, the depot-rooted route visits a node in $A$.
Because the depot is outside $A$ and the route starts and ends at the depot,
it must enter $A$ at least once and leave $A$ at least once.  These two arcs
contribute at least two to $\delta_k(A)$.
\end{proof}

At a relaxation callback, stations with positive fractional visit values are
ranked.  Candidate subsets have size at most four, at most 50 cuts are added
per fixed-interval solve, and only rows violated by more than $10^{-6}$ are
submitted.  The relaxation vector is therefore a separation heuristic only;
the proof above, not the sampled vector, establishes validity.  No candidate
selection rule reads an instance name, seed, path, archived result, or known
objective.

The implementation records candidate count, violated count, added count,
maximum and mean violation, and cut counts by subset size.  It also exports the
first relaxation vector for diagnostic fractionality analysis.  This vector
never enters the certificate ledger.

\subsection{Solver boundary and exact closure}
CPLEX best bounds are admissible only for the original fixed-interval compact
model with paper-safe rows.  Integer infeasibility closes a no-improver leaf;
integer optimality plus independent route reconstruction may update a same-run
upper bound according to the sealed-run policy.  A finite time-limit best
bound may improve the ledger only when its scope and validity flags identify
the original fixed interval.  Wrapper-only heartbeats and diagnostic vectors
cannot close a leaf.

The official plain benchmark uses the current binary-expansion compact MILP
and CPLEX with the same single-thread policy.  It does not use the frontier,
tailored static rows, or route callback and is never imported into the tailored
certificate.  Standard branch-and-cut validity follows the requirement that
user cuts retain every original integer feasible point \cite{wolsey1998}.
