\section{Certificate Ledger and Evidence Boundary}
Each final Gini leaf records its interval, lower bound, closure basis, scope,
solver status, and whether exact coverage remains.  A global certificate is
written only when the verified incumbent exists, the improving domain is
covered, every final leaf is validly closed, no invalid interval or open node
remains, and the aggregate lower bound reaches the incumbent within tolerance.

The admissible sources are valid frontier/relaxation bounds, exact original
fixed-interval infeasibility, valid original fixed-interval CPLEX best bounds,
and empty-domain proofs.  Same-run heuristic routes are upper bounds only and
must pass independent route, load, inventory, Gini, penalty, and objective
verification.

The tested commands disable archive scanning, imported known or external
incumbents, focus-only closure, route-subset enumeration, manual fixed-interval
upper-bound import, and the optional S ledger.  Plain CPLEX rows, callback
vectors, telemetry, and any diagnostic model are benchmark or diagnostic
artifacts only.  The full-frontier merge does not read them.

Audits enforce these rules at three levels.  Source audits reject
instance-specific algorithm branches.  Per-JSON certificate audits check
scope, coverage, verifier status, and forbidden-source flags.  Package audits
require fresh-run provenance and reject summary or raw-result pointers to a
different result round.  Thread and objective-convention audits ensure that
tailored and benchmark rows use one CPLEX thread and the same $\lambda$,
duration, inventory, target, Gini, and penalty conventions.

\begin{corollary}[No diagnostic contamination]
If the certificate-source, package-locality, and paper-strict audits pass, and
the ledger hypotheses of the full-frontier theorem hold, no benchmark or
diagnostic row is needed to establish the reported optimum.
\end{corollary}
\begin{proof}
The allowed closure-source enumeration is exhaustive in the ledger audit.
Every disallowed source is either absent or explicitly marked noncontributing;
the audit rejects any contrary row.  The remaining accepted sources are those
used in the theorem.
\end{proof}
