\section{Computational Protocol}
The fresh evaluation uses source commit recorded in the package manifest and
stores every raw JSON, child interval model, progress trace, solver log, and
summary under one result directory.  No prior raw row or fixed-interval
incumbent is copied.  Tailored and official plain CPLEX rows both use one
solver thread.  A generic wall-budget policy allocates 68\% to frontier
construction, 28\% to exact interval leaves, and 4\% to model and wrapper
overhead; the benchmark receives the matched nominal wall budget directly.
Runs longer than three hours emit five-minute progress and process-tree memory
samples.

The controls are V12/M1, V12/M2, \path{tight_T_seed3101}, and
\path{high_imbalance_seed3202}.  The hard set is
\path{moderate_seed3301}, \path{moderate_seed3302},
\path{high_imbalance_seed3201}, and \path{tight_T_seed3102}.
Comparisons use matched nominal budgets and distinguish certified runtime from
noncertified bound/gap evidence.  Fixed-interval diagnostics are not compared
as if they were full-frontier runs.

\IfFileExists{generated_results.tex}{\input{generated_results}}{
\paragraph{Result table.} The package aggregation step generates
\texttt{generated\_results.tex} from fresh selected rows.}

In this matrix the tailored framework certifies all four controls, although
the required budget is not uniformly lower than plain CPLEX.  None of the
four hard rows is certified.  The long hard rows expose both valid open-ledger
progress and wrapper/native-exit failures; these are reported rather than
converted into closures.  Size-four callback rows are observed, confirming
that the selected separator is active, but this experiment has no matched
no-callback tailored arm.  It therefore does not identify the callback's
causal contribution separately from the rest of the tailored framework.
The mixed hard-row comparison supports continued experimental use of the
profile, not a claim of uniform dominance or a universal default policy.

The computational conclusions are deliberately limited to this matrix.  A
certified control supports stability; an open hard row reports its valid
ledger bound and blocker.  A route-profile advantage requires either an exact
certificate absent from the matched benchmark or a clearly stronger valid
same-budget bound trajectory.  No claim is inferred from callback row count
alone.
